Quantum Mechanics Assignment Nine

### 2016-04-05 quantum/latex/a9.tex
last minute adds.

>>>>

\section{Problem three: p-state transitions}
Here we use McIntyre (14.84) to find which quantum states can be reached from the p-state.
From 14.84 we grab the following statement:\\

\begin{align*}
  \sqrt{\frac{3(2\ell_i+1)}{4\pi(2\ell_f+1)}}\bra{\ell_i,1,m_i,m}\ket{\ell_f,m_f}
  \bra{\ell_i,1,0,0}\ket{\ell_f,0}\\
\end{align*}

We then plug in values for $\ell_i$ and find the CB coefficients that lead to nonzero
values for this expression.\\

By looking at the CB table, we can see that for the $m_1=0,m_2=0$ RHS braket, the only
possible coupled basis J values are 0,2. This means that $\ell_f=0,2$, so by this we
can see that $\ell_i\neq\ell_f$. This is the only selection rule imposed by the RHS
braket, so we ignore it and look instead at the left braket:\\

\begin{align*}
  \bra{1,1,1,1}\ket{\ell_f,m_f}  \rightarrow\bra{\ell_f,m_f}=\braket{2,2}\\
  \bra{1,1,1,0}\ket{\ell_f,m_f}  \rightarrow\bra{\ell_f,m_f}=\braket{2,1}\\
  \bra{1,1,1,-1}\ket{\ell_f,m_f} \rightarrow\bra{\ell_f,m_f}=\braket{2,0},\braket{0,0}\\
  \bra{1,1,0,1}\ket{\ell_f,m_f}  \rightarrow\bra{\ell_f,m_f}=\braket{2,1}\\
  \bra{1,1,0,0}\ket{\ell_f,m_f}  \rightarrow\bra{\ell_f,m_f}=\braket{2,0},\braket{0,0}\\
  \bra{1,1,0,-1}\ket{\ell_f,m_f} \rightarrow\bra{\ell_f,m_f}=\braket{2,-1}\\
  \bra{1,1,-1,1}\ket{\ell_f,m_f} \rightarrow\bra{\ell_f,m_f}=\braket{2,0},\braket{0,0}\\
  \bra{1,1,-1,0}\ket{\ell_f,m_f} \rightarrow\bra{\ell_f,m_f}=\braket{2,-1}\\
  \bra{1,1,-1,-1}\ket{\ell_f,m_f}\rightarrow\bra{\ell_f,m_f}=\braket{2,-2}\\
\end{align*}

From this we can see that the allowed transitions from the p-state are either to the s state
or any of the d-state orbitals.\\

This was a weird problem. It is odd that you can treat the final electron state as
represented in the coupled basis and the original state in the uncoupled basis and get
meaningful results. This is really cool, but an odd usage of the CB table.\\

\section{Problem four: d-state transition}
Here we are given a particular quantum state, $n=4,\ell=3,m_\ell=3$, and asked to find
the possible transitions that can be made from this state. We use the basic selection
rules to answer this question: $\Delta \ell = \pm 1$ and $\Delta m_\ell = 0$.\\

By applying these we see that the following state can be transitioned to:\\

\begin{align*}
  \ket{n, \ell_f,m_{\ell f}} \rightarrow
  &\ket{3, 2, 2}\hspace{1cm}\sigma^-\mbox{ transition}\\
\end{align*}

From this state we can transition to:\\

\begin{align*}
  \ket{3, 2, 2}\rightarrow
  \ket{2, 1, 1}\hspace{1cm}\sigma^-\mbox{ transition}\\
\end{align*}

And then

\begin{align*}
  \ket{2, 1, 1}\rightarrow
  \ket{1, 0, 0}\hspace{1cm}\sigma^-\mbox{ transition}\\
\end{align*}

So we will always be decreasing our z-momentum, so all of our light has $\sigma^-$
polarization. It is interesting that after such complicated arguments we get such
simple selection rules for quantum state transitions.\\
